---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# PyInRpkg

<!-- badges: start -->
<!-- badges: end -->

The goal of PyInRpkg is to play with different ways of wrapping python in an R package to make it feel seamless and also be robust to both auto-managing python environments and not stomp on preexisting environments.

## Installation

You can install the development version of PyInRpkg from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("galenholt/py_in_rpkg")
```
I highly recommend doing this into a `renv`-managed R project to sandbox the testing.

## Background

We want the package to access python functions in modules in `inst/python`, but not expose a bunch of python objects in the R session. To do that they've been wrapped, which can be done in various ways (see the files in `R`).

We also want the package to handle pre-existing python environments (including those in active use in-session), as well as auto-build python and dependencies, which is handled with some code in `R/zzz.R` and a `Config/reticulate` entry in the `DESCRIPTION`.

In general, this follows the suggestions in the [reticulate docs](https://rstudio.github.io/reticulate/articles/python_dependencies.html), I just wanted to get a better sense of how things worked and test different configurations.

In-detail stepthroughs of the thinking and testing to figure these approaches out is at [my website](https://galenholt.github.io/RpyEnvs/py_in_rpkg.html)

## Use

In a bare R project or one with a `.venv` (and I assume other python envs) install as above, then run 

```{r example}
library(PyInRpkg)
```

The R wrappers of the python functions are available, but there are no obvious python objects cluttering up the environment. And in fact, it has not yet loaded python at all.

As soon as we try to use a function, it will auto-configure python, ranging from a full install of python if none is available to doing nothing if python and all dependencies are already present. It does this whether or not the specific function has dependencies.

```{r adder}
adder_wrap(4,5)
```

Functions that depend on packages (e.g. pandas) work too.

```{r}
wrap_min(iris, 'Species') 
wrap_mean(iris, 'Species')
```

Note that the *first* time functions are run they can be slow, but are fast thereafter.

